\Title

### 2013-02-11 synmut.tex
discussion reshuffled

@@

A functional significance of the insulating RNA structure stems between the
hyper-variable loops has been proposed
previously~\citep{watts_architecture_2009, sanjuan_interplay_2011}.
Conserved RNA structures exist in different parts of the HIV genome (several are
known), but our analysis requires longitudinal data, most of which focuses
on \env{}.
Furthermore, we only expect to see deleterious synonymous
mutations at high frequencies if they are linked to repeated adaptive
substitions. Otherwise, they remain at low frequencies and can only be observed
by deep sequencing methods.

The fixation and extinction times and probabilities represent a rich and simple
summary statistics useful to characterize longitudinal sequence data and compare
to models via computer simulations.
A similar method has been recently used in a longitudinal study of
influenza evolution~\citep{strelkowa_clonal_2012}. The propagators suggested in
that paper, however, represent ratios between nonsynonymous and synonymous mutations.
The latter is used as an approximately neutral control and this method can
therefore not be used to investigate synonymous changes themselves.

>>>>

The fixation and extinction times and probabilities represent a rich and simple
summary statistics useful to characterize longitudinal sequence data and compare
to models via computer simulations. Although several recent publications have
pointed out that synonymous changes in HIV can be
deleterious~\citep{lemey_synonymous_2007, ngandu_extensive_2008}, our analysis
is much more direct, as we rely neither on phylogenetic tree reconstruction nor
on macroevolutionary models. A method that is similar to ours {\it in spiritu}
has been recently used in a longitudinal study of influenza
evolution~\citep{strelkowa_clonal_2012}. The propagators suggested in that
article, however, represent ratios between nonsynonymous and synonymous
mutations.  The latter is used as an approximately neutral control; this method
can therefore not be used to investigate synonymous changes themselves. A
functional significance of the insulating RNA structure stems between the
hyper-variable loops has also been proposed
previously~\citep{watts_architecture_2009, sanjuan_interplay_2011} and conserved
RNA structures exist in different parts of the HIV
genome~\citep{ngandu_extensive_2008}. Our analysis is nonetheless much more
direct and is able to quantify the fitness effect of RNA structure within single
infections. The main limitations of our approach are the following: (i) it
requires longitudinal data, most of which focuses on \env{} only; and (ii) we
only expect to see deleterious synonymous mutations at high frequencies if they
are linked to repeated adaptive substitions. Otherwise, most synonymous
polymorphisms remain at low frequencies and can only be observed by deep
sequencing methods. We indeed plan to exploit next-generation sequencing
technologies to assess synonymous diversity at the level of rare variants in
future studies.

@@
